Free Probability Analysis for Reflect Surface aided Massive MIMO systems

### 2020-06-04 documentation/Paper/paper.tex
working on free probability background of paper

\chapter{Introduction}

\section{Notation}
The following notation is used throughout the remainder. 
$E\{x\}$ is the expected value of a random variable $x$.
Vectors are denoted by bold font, lower-case letters ($\vx$) and are assumed to be column vectors.
Matrices are denoted by bold font, upper-case letters ($\mX$). The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
$\mathbf{I}$ denotes the identity matrix.

@@

\item 
	Brief introduction to free probability.
	Use the example of a degenerate matrix with rank 1 being multiplied by a random matrix. 
	In this case the resulting pdf is not just the convolution of the original distributions. 
	In order to introduce the tools that will be used in the following, an example is now given to
	exhibit this case in which classic probability (commutative probability) no longer holds.
	Consider the Gaussian i.i.d. matrix $\mH$ with $ h_{i,j} \in \CN$ and the diagonal matrix $\mP$ with 
	$p_{i,i} = 1$ for $i = 1 \cdots K$ with $K \leq N$ and both matrices of dimension $N \times N$. 
	We are interested the AED for $\mP\mH \mH^H \mP^H$ and immediately note  that the matrices
	$ \mP^H \mP \mH \mH^H = \mP \mH \mH^H$ have the same non-zero eigenvalues (show why!). 
	Given that both $ \mP$ and  $\mH \mH^H$ have converging AEDs as $N \rightarrow \infty$
	classical probability would tell us that the AED for $\mP \mH \mH^H$ is the convolution of the two AEDs. 
	This, however, is not the case non-commutative distributions. This can be seen without the tools of free probability by noting
	that the resulting product has $\Rank = \min[\Rank(\mP), \Rank(\mH \mH^H)]$. In the case that $\Rank(\mP) = 1$
	the resulting product must have rank 1 (i.e a single non-zero eigenvalue) with corresponding AED
	$p_{\mP}(x) = \frac{1}{N-1}\delta(x)  + \frac{1}{N}\delta(x-1)$. Following classical probability, the convolution of 
	a general AED with $p_{\mP}(x)$, however, can be continuous. 
	\par 
	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.

@@

@@

>>>>

\chapter{Introduction}
Wireless communication has progressively moved to higher frequencies in recent standards in order to utilize large bands of unoccupied resource
and to enable increased antenna density (CITE). One hurdle to higher frequencies is increased radio frequency energy (power) attenuation (CITE). 
In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
and supplementing the transmission power using relays (CITE). Two key costs to deploying relays are the radio frequency chains used for receiving and transmitter, and the power used for transmission. (Would be nice to cite some actual costs here). As the power attenuation problem has been increased by X-Fold moving from freq1 to freq2, new solutions have been considered. One solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves. In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
\par
The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
associated with a relay. The absence of an rf-chain in the RIS offers additional benefits over relays. First, reflection does not introduce the noise associated with the rf-chain. Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum. Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on a relay. 
The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is proportional to $N^2$ (? Does this hold if we remove the ability to choose the angle?).
For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
Example of how these surfaces are implemented (Cite).
Discuss how it has be looked at from MISO and MIMO perspective.
\par
One area that has not be investigated, (at least up to the author's knowledge) is the impact of the RIS for 
asymptotically large MIMO systems. In the following, the impact of the RIS on a general MIMO system is explored.

\section{Notation}
The following notation is used throughout the remainder. 
$\Expect[x]$ is the expected value of a random variable $x$.
Bold font, lower-case letters ($\vx$) and upper-case letters ($\mX$) denote vector and matrices respectively and all vectors are assumed to be column vectors unless noted otherwise. 
The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
$\mI_N$ denotes the identity matrix of dimension $N$.

@@

\item 
	A brief introduction to free probability.
	In order to introduce the tools that will be used in the following, an example of non-commutative probability now given to
	exhibit a case in which the tools of classic probability are not sufficient for finding a probability density function (pdf).
	\par 
	Free probability is the counterpart to classical probability for the case in which the random variables (in this case random matrices) do 
	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables (It is important to note here
	 that by changing this nature of the random variables from commutative to non-commutive (is this embedded in kolmogorov's axioms?) 
	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context 
	but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.

@@

\item 
	Give proofs in an appendix?

@@

\chapter{Appendix}
Proofs
\begin{enumerate}
\item
	Prove freeness of the relevant points. 
\item 
	rotations have same non-zero eigenvalues
\end{enumerate}
